Predicting the Quality of CDRH3 Antibody Loop Structural Models

### 2022-10-05 paper.tex
Fixed all full-stops commas and dashes

Antibodies consist of a heavy and a light chain, which are linked by
disulphide bonds. The N-terminal domain of each makes up the variable
fragment (Fv), which contains the complementarity determining regions
(CDRs). The antigen binding site is composed of six CDRs, also known
as hypervariable loops. All except one of these loops can be clustered
into a limited number of canonical structures. Therefore, modelling
these loops with adequate accuracy is commonly
achievable\cite{North2011,Weitzner2015}. However, the CDR loop 3 of
the 
heavy chain (CDR-H3) has a far greater sequence variability due to the
processes of V(D)J recombination and somatic hyper‐mutation and its
structure has remained unclassifiable\cite{Finn2016}. The variety in
structure is so great, that its structural diversity is remarkable
even compared to other protein loops\cite{Regep2017}. It was found
that over 75\% of CDR-H3 loops do not have a sub-Angstrom non-antibody
structural neighbour, as well as that 30\% of CDR-H3 loops have a
completely unique structure compared with under 3\% for all other
loops on average\cite{Regep2017}.

@@

According to the Kabat definition, the CDR-H3 loop is made up of the
residues 95-105 (Kabat numbering scheme\cite{Kabat1992}) in the heavy
chain, with a potential insertion site at position 100. The
possibility of such an insertion of a varying number of residues leads
to a large range of loop lengths, with bovine antibodies being
exceptionally long (Figure~\ref{fig:loopdist}).

\begin{figure}
  \includegraphics[scale=1]{loopdist.eps}
  \centering
  \caption { Distribution of CDR-H3 loop lengths in mouse (left),
    human (centre) and cow (right). Mouse and human antibody CDR-H3s
    have a unimodal, yet almost normal distribution with a range of
    ca. 4-28 and 4-38 respectively. The length cut-off for CDR-H3s
    depicted above is 30 amino acids. Bovine antibodies with
    significantly longer CDR-H3 loops than depicted here exist,
    reaching a length of 67 amino acids and
    above\protect\cite{Wong2015}.}  
  \label{fig:loopdist}
\end{figure}

@@

Using these mentioned modelling methods, framework regions can
generally be predicted with great accuracy (with better than 1\AA\
RMSD\cite{Almagro2014}), as one can often find a very similar
structure for the homology modelling process.  However, the CDR loops
are not as easily predicted due to their great diversity. If the
canonical conformation of CDR loops 1-5 can be identified, they too
can be modelled rather well, often within 1\AA\ RMSD, for CDR-H3 loops
the average is usually above 3\AA\cite{Almagro2011}.

@@

The target data (i.e.\ RMSD values) are then transformed from numerical
values to nominal values so that they can be used for
classification. In order to define these nominal categories, the total
RMSD range must be divided into categories. This is done either by
creating uniform classes i.e.\ 1-2~\AA, 2-3~\AA\ (the optimal size of which
must be determined), etc. or by creating balanced classes. When
creating balanced classes, the upper and lower thresholds of a
category are chosen in such a way that each class contains an equal
number of instances. This approach is chosen to counteract the
skewness of the RMSD distribution. However, this was found to
negatively affect the final model's predictive power. Therefore,
uniform classes were used.

@@

The classifier predicts whether a model has an RMSD of below 2~\AA,
between 2-4~\AA, or above 4~\AA. These cut-off values were selected
based on the observation that abYmod generally produces a model with
RMSD below 4~\AA. Incorrectly modelled structures (Figure~\ref{fig:AMA}) may
be identified by screening for structures estimated to have an RMSD
above 4~\AA. If a very high-quality model is needed one should also
exclude models above 2~\AA.

@@

\subsection{Computing}
All machine learning and, feature selection and hyperparameter
optimization algorithms were implemented in Python. The Scikit-learn
library was used for training models, the
Yellowbrick\cite{Bengfort2021} library was utilized for
visualization. All code is available at
https://github.com/LilianDenzler/qualiloop

@@

\subsubsection{Physiochemical Feature Encoding}
In a paper by L.~Nanni and A.~Lumini a new encoding technique is
presented which was developed for machine learning classifiers. Many
physiochemical properties are calculated and transformed using a
non-linear Fisher transform for dimensionality reduction.  A vector of
length 19 is produced for each amino acid34.  In a paper on designing
a neural network for predicting the packing angle of the light and
heavy variable chain of an antibodies, A.C.R.~Martin and
K.R.~Abhinandan introduce an encoding method that produces a
four-dimensional physiochemical feature vector25. The amino acid
properties used are 1) the total number of side-chain atoms, 2) the
number of side-chain atoms in the shortest path from \ca\ to the most
distal atom, 3) the Eisenberg consensus
hydrophobicity\cite{Eisenberg1982}, 4) the charge.

@@

The selection of features incorporated in the training set seemed to
be most important for effective learning. A multitude of methods were
tested. No one-fit-for-all method for the different models could be
found. However, for our top-layer classifier in our final model
recursive feature elimination worked best.

>>>>

Antibodies consist of a heavy and a light chain, which are linked by
disulphide bonds. The N-terminal domain of each makes up the variable
fragment (Fv), which contains the complementarity determining regions
(CDRs). The antigen binding site is composed of six CDRs, also known
as hypervariable loops. All except one of these loops can be clustered
into a limited number of canonical structures. Therefore, modelling
these loops with adequate accuracy is commonly
achievable\cite{North2011,Weitzner2015}. However, the CDR loop 3 of
the 
heavy chain (CDR-H3) has a far greater sequence variability due to the
processes of V(D)J recombination and somatic hyper‐mutation and its
structure has remained unclassifiable\cite{Finn2016}. The variety in
structure is so great, that its structural diversity is remarkable
even compared to other protein loops\cite{Regep2017}. It was found
that over 75\% of CDR-H3 loops do not have a sub-{\AA}ngstr\"{o}m non-antibody
structural neighbour, as well as that 30\% of CDR-H3 loops have a
completely unique structure compared with under 3\% for all other
loops on average\cite{Regep2017}.

@@

According to the Kabat definition, the CDR-H3 loop is made up of the
residues 95--105 (Kabat numbering scheme\cite{Kabat1992}) in the heavy
chain, with a potential insertion site at position 100. The
possibility of such an insertion of a varying number of residues leads
to a large range of loop lengths, with bovine antibodies being
exceptionally long (Figure~\ref{fig:loopdist}).

\begin{figure}
  \includegraphics[scale=1]{loopdist.eps}
  \centering
  \caption { Distribution of CDR-H3 loop lengths in mouse (left),
    human (centre) and cow (right). Mouse and human antibody CDR-H3s
    have a unimodal, yet almost normal distribution with a range of
    $\sim$4--28 and 4--38 respectively. The length cut-off for CDR-H3s
    depicted above is 30 amino acids. Bovine antibodies with
    significantly longer CDR-H3 loops than depicted here exist,
    reaching a length of 67 amino acids and
    above\protect\cite{Wong2015}.}  
  \label{fig:loopdist}
\end{figure}

@@

Using these mentioned modelling methods, framework regions can
generally be predicted with great accuracy (with better than 1\AA\
RMSD\cite{Almagro2014}), as one can often find a very similar
structure for the homology modelling process.  However, the CDR loops
are not as easily predicted due to their great diversity. If the
canonical conformation of CDR loops 1--5 can be identified, they too
can be modelled rather well, often within 1\AA\ RMSD, for CDR-H3 loops
the average is usually above 3\AA\cite{Almagro2011}.

@@

The target data (i.e.\ RMSD values) are then transformed from numerical
values to nominal values so that they can be used for
classification. In order to define these nominal categories, the total
RMSD range must be divided into categories. This is done either by
creating uniform classes i.e.\ 1--2~\AA, 2--3~\AA\ (the optimal size of which
must be determined), etc.\ or by creating balanced classes. When
creating balanced classes, the upper and lower thresholds of a
category are chosen in such a way that each class contains an equal
number of instances. This approach is chosen to counteract the
skewness of the RMSD distribution. However, this was found to
negatively affect the final model's predictive power. Therefore,
uniform classes were used.

@@

The classifier predicts whether a model has an RMSD of below 2~\AA,
between 2--4~\AA, or above 4~\AA. These cutoff values were selected
based on the observation that abYmod generally produces a model with
RMSD below 4~\AA. Incorrectly modelled structures (Figure~\ref{fig:AMA}) may
be identified by screening for structures estimated to have an RMSD
above 4~\AA. If a very high-quality model is needed one should also
exclude models above 2~\AA.

@@

\subsection{Computing}
All machine learning and feature selection and hyperparameter
optimization algorithms were implemented in Python. The Scikit-learn
library was used for training models, the
Yellowbrick\cite{Bengfort2021} library was utilized for
visualization. All code is available at
https://github.com/LilianDenzler/qualiloop

@@

\subsubsection{Physiochemical Feature Encoding}
In a paper by L.~Nanni and A.~Lumini a new encoding technique is
presented which was developed for machine learning classifiers. Many
physiochemical properties are calculated and transformed using a
non-linear Fisher transform for dimensionality reduction.  A vector of
length 19 is produced for each amino acid34.  In a paper on designing
a neural network for predicting the packing angle of the light and
heavy variable chain of an antibodies, A.C.R.~Martin and
K.R.~Abhinandan introduce an encoding method that produces a
four-dimensional physiochemical feature vector25. The amino acid
properties used are 1) the total number of sidechain atoms, 2) the
number of sidechain atoms in the shortest path from \ca\ to the most
distal atom, 3) the Eisenberg consensus
hydrophobicity\cite{Eisenberg1982}, 4) the charge.

@@

The selection of features incorporated in the training set seemed to
be most important for effective learning. A multitude of methods were
tested. No one fit-for-all method for the different models could be
found. However, for our top-layer classifier in our final model
recursive feature elimination worked best.
